---
title: "Hackathon Analysis"
output:
  pdf_document: 
    fig_height: 6
    fig_width: 8
    toc: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analyse 
### Project composition and functions 

### 1) RNA-binding proteins (filter_ATtRACT)

```{r, echo=FALSE, message=FALSE, results="hide"}
#Visualize the entropy
source("human-RBP-analysis/01_filter_ATtRACT.R")
```

### 2) Genome

### 3) Annotated sites
We load the gff file. We can visualize the head of the file.  
```{r, echo = FALSE, out.width = "100%",fig.align = "center"}
library(png)
img1_path <- "images/gff_head.png"
img1 <- readPNG(img1_path, native = TRUE, info = TRUE)
grid::grid.raster(img1)
```

We can also visualize the head of the sites file. 
```{r, echo = FALSE, out.width = "100%",fig.align = "center"}
library(png)
img2_path <- "images/sites_head.png"
img2 <- readPNG(img2_path, native = TRUE, info = TRUE)
grid::grid.raster(img2)
```

The new file (*annoted_sites*) is more complete. 
```{r, echo = FALSE, out.width = "100%",fig.align = "center"}
library(png)
img3_path <- "images/annotated_sites_head.png"
img3 <- readPNG(img3_path, native = TRUE, info = TRUE)
grid::grid.raster(img3)
```

### 4) Simulated genomes 

```{r, echo=FALSE, message=FALSE, results="hide"}
source("human-RBP-analysis/04_simulate_genomes.R")
```

```{r,echo=FALSE, message=FALSE}
load("D:/Columbia/Genomique/Final Project/ecbm-e4060-covid-interactions/hackathon_published/output50/sim_genomes.RData")
head(data.frame(sim_genomes))
```

### 5) Find the binding sites of the simulated genomes

### 6) Enrichment analysis 

### 7) Filter hits

```{r, echo = FALSE, out.width = "100%",fig.align = "center"}
library(png)
img6_path <- "images/sites_counts_paper.png"
img6 <- readPNG(img6_path, native = TRUE, info = TRUE)
grid::grid.raster(img6)
```

#### a) **50 simulations** 

```{r, echo=FALSE, message=FALSE}
library(kableExtra)
library(knitr)

load("D:/Columbia/Genomique/Final Project/ecbm-e4060-covid-interactions/hackathon_published/output50/genome_hits.RData")
knitr::kable(genome_hits, caption="genome_hits") %>% kable_styling(bootstrap_options = c("striped", "hover"),latex_options = "hold_position",font_size = 10 ) %>% 
  column_spec(1, bold = T, color = "white", background = "#66C2A5") %>% 
  row_spec(0, bold = T, color = "white", background = "#66C2A5")

load("D:/Columbia/Genomique/Final Project/ecbm-e4060-covid-interactions/hackathon_published/output50/f_utr_hits.RData")
knitr::kable(f_utr_hits, caption="5’ UTR sequences") %>% kable_styling(bootstrap_options = c("striped", "hover"),latex_options = "hold_position",font_size = 10 ) %>% 
  column_spec(1, bold = T, color = "white", background = "#66C2A5") %>% 
  row_spec(0, bold = T, color = "white", background = "#66C2A5")

load("D:/Columbia/Genomique/Final Project/ecbm-e4060-covid-interactions/hackathon_published/output50/t_utr_hits.RData")
knitr::kable(t_utr_hits, caption="3’ UTR sequences") %>% kable_styling(bootstrap_options = c("striped", "hover"),latex_options = "hold_position",font_size = 10 ) %>% 
  column_spec(1, bold = T, color = "white", background = "#66C2A5") %>% 
  row_spec(0, bold = T, color = "white", background = "#66C2A5")

load("D:/Columbia/Genomique/Final Project/ecbm-e4060-covid-interactions/hackathon_published/output50/genome_neg_hits.RData")
knitr::kable(genome_neg_hits, caption="Genome negative hits") %>% kable_styling(bootstrap_options = c("striped", "hover"),latex_options = "hold_position",font_size = 10 ) %>% 
  column_spec(1, bold = T, color = "white", background = "#66C2A5") %>% 
  row_spec(0, bold = T, color = "white", background = "#66C2A5")
```

```{r}
load("D:/Columbia/Genomique/Final Project/ecbm-e4060-covid-interactions/hackathon_published/output5000/genome_hits.RData")
genome_hits<-genome_hits[,c(-2,-4,-5,-6,-8)]
knitr::kable(genome_hits, caption="genome_hits") %>% kable_styling(bootstrap_options = c("striped", "hover"),latex_options = "hold_position",font_size = 10 ) %>% 
  column_spec(1, bold = T, color = "white", background = "#66C2A5") %>% 
  row_spec(0, bold = T, color = "white", background = "#66C2A5")
```


```{r}
load("D:/Columbia/Genomique/Final Project/ecbm-e4060-covid-interactions/hackathon_published/output5000/f_utr_hits.RData")
f_utr_hits<-f_utr_hits[,c(-2,-4,-5,-6,-8)]
knitr::kable(f_utr_hits, caption="5’ UTR sequences") %>% kable_styling(bootstrap_options = c("striped", "hover"),latex_options = "hold_position",font_size = 10 ) %>% 
  column_spec(1, bold = T, color = "white", background = "#66C2A5") %>% 
  row_spec(0, bold = T, color = "white", background = "#66C2A5")

load("D:/Columbia/Genomique/Final Project/ecbm-e4060-covid-interactions/hackathon_published/output5000/t_utr_hits.RData")
t_utr_hits<-t_utr_hits[,c(-2,-4,-5,-6,-8)]
knitr::kable(t_utr_hits, caption="3’ UTR sequences") %>% kable_styling(bootstrap_options = c("striped", "hover"),latex_options = "hold_position",font_size = 10 ) %>% 
  column_spec(1, bold = T, color = "white", background = "#66C2A5") %>% 
  row_spec(0, bold = T, color = "white", background = "#66C2A5")

load("D:/Columbia/Genomique/Final Project/ecbm-e4060-covid-interactions/hackathon_published/output5000/genome_neg_hits.RData")
genome_neg_hits<-genome_neg_hits[,c(-2,-4,-5,-6,-8)]
knitr::kable(genome_neg_hits, caption="Genome negative hits") %>% kable_styling(bootstrap_options = c("striped", "hover"),latex_options = "hold_position",font_size = 10 ) %>% 
  column_spec(1, bold = T, color = "white", background = "#66C2A5") %>% 
  row_spec(0, bold = T, color = "white", background = "#66C2A5")
```


**Genome hits**

**Negative genome hits**

**5’ and 3’ UTR sequences**

```{r, echo=FALSE, message=FALSE}
#par(mfrow = c(1,2),mar=c(5, 3, 1, 1)) 
library("RColorBrewer")
colors1 <- brewer.pal(n = 11, name = "Set3")
colors2 <- brewer.pal(n = 9, name = "Set2")
colors<-c(colors1, colors2)

barplot(genome_hits$pval, main = "Pvalue for the 19 genome hits",las=2,horiz=F, col = colors, cex.main = 1, cex.names = 0.7, ylab="number of sites", xlab="Gene_name", names.arg=genome_hits$Gene_name)

barplot(genome_hits$N, main = "Number of sites for the 19 genome hits",las=2,horiz=F, col = colors, cex.main = 1, cex.names = 0.7, ylab="number of sites", xlab="Gene_name", names.arg=genome_hits$Gene_name)


```

```{r, echo=FALSE, message=FALSE}
par(mfrow = c(1,2),mar=c(4.5, 4, 1, 1)) 
barplot(genome_neg_hits$pval, main = "Number of sites for negative genome hits",las=2,horiz=F, col = colors, cex.main = 1, cex.names = 0.7, ylab="number of sites", xlab="Gene_name", names.arg=genome_neg_hits$Gene_name)

barplot(genome_neg_hits$N, main = "Number of sites for negative genome hits",las=2,horiz=F, col = colors, cex.main = 1, cex.names = 0.7, ylab="number of sites", xlab="Gene_name", names.arg=genome_neg_hits$Gene_name)


```

```{r, echo=FALSE, message=FALSE}
par(mfrow = c(1,2),mar=c(6, 4, 1, 1)) 
barplot(t_utr_hits$pval, main = "Number of sites for 3' sequences",las=2,horiz=F, col = colors, cex.main = 1, cex.names = 0.7, ylab="number of sites", xlab="Gene_name", names.arg=t_utr_hits$Gene_name)

barplot(t_utr_hits$N, main = "Number of sites for 3' sequences",las=2,horiz=F, col = colors, cex.main = 1, cex.names = 0.7, ylab="number of sites", xlab="Gene_name", names.arg=t_utr_hits$Gene_name)

```

# Results 

```{r}
# Import requirements

library(data.table)
source("human-RBP-analysis/rbp_functions.R")

################################################

# Compare hits in whole genome

load("output50/genome_hits.RData")
sars2 = genome_hits

load("output50/MN996532.2/genome_hits.RData")
bat = genome_hits

load("output50/NC_004718.3/genome_hits.RData")
sars = genome_hits

diff_hits_sars2_bat=setdiff(sars2[, Gene_name], bat[, Gene_name]) 
paper_hits_sars2_bat="character(0)"
diff_hits_sars2_sars=setdiff(sars2[, Gene_name], sars[, Gene_name]) #
paper_hits_sars2_sars=c("PABPC1", "ZNF638" ,"PABPC5", "CELF2",  "PABPC3", "YBX1")
################################################

# Compare hits in 3'UTR
load("output50/t_utr_hits.RData")
sars2 = t_utr_hits

load("output50/MN996532.2/t_utr_hits.RData")
bat = t_utr_hits

load("output50/NC_004718.3/t_utr_hits.RData")
sars = t_utr_hits

diff_3_sars2_bat=setdiff(sars2[, Gene_name], bat[, Gene_name]) 
paper_3_sars2_bat="character(0)"
diff_3_sars2_sars=setdiff(sars2[, Gene_name], sars[, Gene_name]) 
paper_3_sars2_sars="character(0)"

###############################################
# Compare hits in 5'UTR

load("output50/f_utr_hits.RData")
sars2 = f_utr_hits

load("output50/MN996532.2/f_utr_hits.RData")
bat = f_utr_hits

load("output50/NC_004718.3/f_utr_hits.RData")
sars = f_utr_hits

diff_5_sars2_bat=setdiff(sars2[, Gene_name], bat[, Gene_name]) #character(0)
paper_5_sars2_bat="character(0)"
diff_5_sars2_sars=setdiff(sars2[, Gene_name], sars[, Gene_name]) 
paper_5_sars2_sars="character(0)"

###############################################

# Compare hits in - strand
load("output50/genome_neg_hits.RData")
sars2 = genome_neg_hits

load("output50/MN996532.2/genome_neg_hits.RData")
bat = genome_neg_hits

load("output50/NC_004718.3/genome_neg_hits.RData")
sars = genome_neg_hits

diff_neg_sars2_bat=setdiff(sars2[, Gene_name], bat[, Gene_name])
paper_neg_sars2_bat=c("ELAVL1",  "HNRNPDL" ,"RBFOX1")
diff_neg_sars2_sars=setdiff(sars2[, Gene_name], sars[, Gene_name]) 
paper_neg_sars2_sars=c("SRSF3" , "TIA1"  , "ELAVL1" ,"FUS" )  

#############################################
a=data.frame(diff_neg=paste(diff_neg_sars2_sars,collapse = " "),paper_neg=paste(paper_neg_sars2_sars,collapse = " "))
knitr::kable(a, caption = "Conclusion, TARGETS") 
```



